|

### 2021-11-23 03-Chp-Systematic-Review-Methods.Rmd
Citations fixed!

<!-- TODO Extra studies to include: -->
<!-- * Larsson - 10.1136/bmj.j5375 [@larsson2017b] -->

@@

I systematically searched several electronic bibliographic databases to identify potentially relevant entries (hereafter referred to as "records"). The following databases were searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection. As the contents of the Web of Science Core Collection can vary by institution,[@gusenbauer2020a] the specific databases and date ranges for each database searched via this platform are listed in Appendix \@ref(appendix-wos-databases). The search strategy used in each database was developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE)[@lefebvre2019searching] terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy included terms related to lipids, lipid modifying treatments, and dementia, and was designed for MEDLINE before being adapted for use in the other bibliography databases listed. A high-level outline of the strategy is presented in the Table \@ref(tab:searchOverview-table) below and the full search strategies for each database are presented in Appendix \@ref(appendix-search-strategy). <!-- TODO Need to actually attach each search strategy. Should be able to loop through search strategy results. -->

@@

Of note, this review did not include the commonly cited PROSPER RCT, which examined the effect of pravastatin on CVD risk, reporting on cognitive outcomes as one of several secondary outcomes.[@shepherd2002a] While widely cited in relation to the effect of statins on dementia risk and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial reported solely on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up. Though an useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using recognised criteria as cognitive tests should feed into a broader diagnostic pathway (see Section \@ref(diagnostic-criteria)). As such, this trial did not met the inclusion criteria for this review.

@@

Risk of bias across the individual results was generally quite high, and the causes and expected directions of these biases are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of statin levels. In many cases, estimates were know to be missing from meta-analyses not at random due to preferential reporting of significant results observed in a number of analysis, leading to the high-risk judgement. These missing estimates were most commonly identified via analysis of conference abstract/final publication pairs.[@yamada2009; @yamada2009a] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the Table.").[@ancelin2013] However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analysis (which provided very weak evidence for an effect), other than increasing the precision of the summary estimate.

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018a; @kuzma2018a] However, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include[@chu2018;@power2015] or used a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

@@

As a further stumbling block, Mendelian randomisation, particularly when using a two-sample summary data design, is a form of analysis that lends itself to multiple exposure-outcome comparisons. This is particularly relevant to the consideration of bias due to missing evidence. As an example, through snowballing and other measures, I identified at least one relevant Mendelian randomisation study that had not been identified by the search strategy.[@larsson2017b] On review of this paper, the search would not have been expected to find it given the absence of any lipid-related keywords in the title and abstract. The study examined the association between lipid fractions with Alzheimer's disease as one of many risk factors for the condition. Studies such as this can introduce bias into a systematic review, as it is commonly only those risk factors that show a statistically significant result that are reported in the abstract and so are captured by the search. This may bias systematic reviews, including this one, as the analysis of multiple risk factors against a single outcome within a single publication becomes more common. These studies are described as "unknown unknown's" in the context of the RoB-ME tool, and are particularly challenging (as opposed to an analysis that was insufficiently reported to be included in the statistical analysis, or the "known unknown's").

@@

One item of particular interest is the attenuation of any effects observed by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the Apoe4 gene region. As covered in the introduction (see Section \@ref(intro-basic-science)), an increasing number of ApoE4 alleles is a major independent risk factor for Alzheimer's disease, and so violates the exclusion restriction criteria (Figure \@ref(fig:mrAssumptions)). In all cases, excluding these variants attenuates the observed effect to the null. A clear example of this is Benn _et al_ (2017), where the ApoE variants were not sufficient identified and excluded and the published paper detailed evidence for a protective effect of LDL-c was identified (`r estimate(0.83, .75,.92,"RR")`).[@benn2017] Following several rapid responses, the data was re-analysed excluding a larger area around ApoE4 which attenuated the finding to the null.[@benn2017a]

@@

Firstly, preprints provide an advance snapshot of the literature, capturing articles that will eventually be published but were not available at the time of the main search. Consider the example of one eligible preprint in this review was initially posted on bioRxiv in July 2019[@andrews2019b] and was subsequently published in 2021 following peer-review.[@andrews2021] Secondly, inclusion of preprints allows for results that may never be formally published to be included in an evidence synthesis exercise as is the case with the second preprint included in this review.[@so2017] Both of these aspects illustrate that, if the aim is to find the current state of the art in the topic area at the time of searching, inclusion of preprints is a necessity.

>>>>

<!-- TODO Extra studies to include: -->
<!-- * Larsson - 10.1136/bmj.j5375 [@larsson2017] -->

@@

I systematically searched several electronic bibliographic databases to identify potentially relevant entries (hereafter referred to as "records"). The following databases were searched from inception onwards: Medline, EMBASE, Psychinfo, Cochrane Central Register of Controlled Trials (CENTRAL), and Web of Science Core Collection. As the contents of the Web of Science Core Collection can vary by institution,[@gusenbauer2020] the specific databases and date ranges for each database searched via this platform are listed in Appendix \@ref(appendix-wos-databases). The search strategy used in each database was developed in an iterative manner using a combination of free text and controlled vocabulary (MeSH/EMTREE)[@lefebvre2019searching] terms to identify studies which have examined the relationship between blood lipids levels and dementia, incorporating input from an information specialist. The strategy included terms related to lipids, lipid modifying treatments, and dementia, and was designed for MEDLINE before being adapted for use in the other bibliography databases listed. A high-level outline of the strategy is presented in the Table \@ref(tab:searchOverview-table) below and the full search strategies for each database are presented in Appendix \@ref(appendix-search-strategy). <!-- TODO Need to actually attach each search strategy. Should be able to loop through search strategy results. -->

@@

Of note, this review did not include the commonly cited PROSPER RCT, which examined the effect of pravastatin on CVD risk, reporting on cognitive outcomes as one of several secondary outcomes.[@shepherd2002] While widely cited in relation to the effect of statins on dementia risk and included in the Cochrane review of RCTs on this topic,[@mcguinness2016] the trial reported solely on the change in a range of cognitive measures (MMSE, Stroop test, Picture-Word Learning test and others) over follow-up. Though an useful indicator of general cognitive decline, it is not equivalent to a dementia diagnosis using recognised criteria as cognitive tests should feed into a broader diagnostic pathway (see Section \@ref(diagnostic-criteria)). As such, this trial did not met the inclusion criteria for this review.

@@

Risk of bias across the individual results was generally quite high, and the causes and expected directions of these biases are discussed in more detail in Chapter \@ref(tri-heading). Of particular interest to this chapter, however, was the high risk of bias due to missing evidence observed for observational studies of statin levels. In many cases, estimates were know to be missing from meta-analyses not at random due to preferential reporting of significant results observed in a number of analysis, leading to the high-risk judgement. These missing estimates were most commonly identified via analysis of conference abstract/final publication pairs.[@yamada2009conf; @yamada2009] In addition, some authors stated outright that non-significant results were not reported (e.g. "The other lipid variables not significantly associated with dementia and Alzheimer’s disease ... were not reported in the Table.").[@ancelin2013] However, as all identified missing results are likely to be non-significant, they would not be expected to have a substantial impact on their respective meta-analysis (which provided very weak evidence for an effect), other than increasing the precision of the summary estimate.

@@

While conducting this review, I identified several previous systematic reviews of this topic.[@chu2018; @yang2020; @muangpaisan2010; @poly2020;@kuzma2018stroke; @kuzma2018risk] However, this review is the first to use established domain based assessments tools (for example, the RoB 2 tool for randomized controlled trials)[@sterne2019] to assess the risk of bias in included studies. The majority of the highly cited reviews on this topic either do not formally consider risk of bias in the observational studies they include[@chu2018;@power2015] or used a non-domain-based assessment tool (e.g. the Newcastle-Ottowa Scale).[@poly2020]

@@

As a further stumbling block, Mendelian randomisation, particularly when using a two-sample summary data design, is a form of analysis that lends itself to multiple exposure-outcome comparisons. This is particularly relevant to the consideration of bias due to missing evidence. As an example, through snowballing and other measures, I identified at least one relevant Mendelian randomisation study that had not been identified by the search strategy.[@larsson2017] On review of this paper, the search would not have been expected to find it given the absence of any lipid-related keywords in the title and abstract. The study examined the association between lipid fractions with Alzheimer's disease as one of many risk factors for the condition. Studies such as this can introduce bias into a systematic review, as it is commonly only those risk factors that show a statistically significant result that are reported in the abstract and so are captured by the search. This may bias systematic reviews, including this one, as the analysis of multiple risk factors against a single outcome within a single publication becomes more common. These studies are described as "unknown unknown's" in the context of the RoB-ME tool, and are particularly challenging (as opposed to an analysis that was insufficiently reported to be included in the statistical analysis, or the "known unknown's").

@@

One item of particular interest is the attenuation of any effects observed by Mendelian randomisation studies following the adjustment for/exclusion of genetic variation in the Apoe4 gene region. As covered in the introduction (see Section \@ref(intro-basic-science)), an increasing number of ApoE4 alleles is a major independent risk factor for Alzheimer's disease, and so violates the exclusion restriction criteria (Figure \@ref(fig:mrAssumptions)). In all cases, excluding these variants attenuates the observed effect to the null. A clear example of this is Benn _et al_ (2017), where the ApoE variants were not sufficient identified and excluded and the published paper detailed evidence for a protective effect of LDL-c was identified (`r estimate(0.83, .75,.92,"RR")`).[@benn2017] Following several rapid responses, the data was re-analysed excluding a larger area around ApoE4 which attenuated the finding to the null.[@benn2017comment]

@@

Firstly, preprints provide an advance snapshot of the literature, capturing articles that will eventually be published but were not available at the time of the main search. Consider the example of one eligible preprint in this review was initially posted on bioRxiv in July 2019[@andrews2019] and was subsequently published in 2021 following peer-review.[@andrews2021] Secondly, inclusion of preprints allows for results that may never be formally published to be included in an evidence synthesis exercise as is the case with the second preprint included in this review.[@so2017] Both of these aspects illustrate that, if the aim is to find the current state of the art in the topic area at the time of searching, inclusion of preprints is a necessity.

### 2021-11-23 05-Chp-CPRD-Analysis.Rmd
End of day:
Citations fixed!

In this analysis, missing data were handled using a multiple imputation approach.[@sterne2009] Variables with missing observations were identified for inclusion in the imputation model. Nominal variables with missing values were modelled using multinomial logistic regression, while continuous variables were modelled using linear regression. As per best practice, all variables used in the analytic model, including the outcome, were included in the imputation model.[@moons2006using] Using the MICE (Multiple Imputation by Chained Equations) command in STATA16, 20 imputed datasets were created.

@@

Using multiple imputation to handle missing data is an alternative to a "complete case" approach,[@pigott2001review] where participants missing any covariate are dropped from the dataset. As a recommended sensitivity analysis,[@hughes2019] I preformed and compared the results of both methods, to investigate the impact of multiple imputation on the results.

@@

Despite observational analyses suggesting a link between statins and muscular pain (as opposed to more serious complications such as myopathy),[@selva-ocallaghan2018] systematic reviews of the adverse events of statin use[@collins2016a] and N-of-1 trials explicitly exploring the association of statin use with muscle pain[@herrett2021] have found little evidence supporting an effect. This suggests that muscular backpain would be suitable for use as a negative control outcome in this analysis. Under this approach, if statin use is found to be associated with muscular backpain in this analysis, this suggests the presence of residual confounding and reduces my confidence in the results for the dementia outcomes.

Similarly, incident ischemic heart disease and Type 2 diabetes were included as a protective and harmful positive control outcome, respectively. The protective effect of lipid-lowering treatment, via statins, on the risk of ischemic heart disease is well-established, [@collins2016a] while there is growing evidence for an increased risk of Type 2 diabetes with statin use.[@collins2016a; @macedo2014; @smit2020] Similar to the negative outcome, if the analysis strategy can reproduce these known associations, this will provide evidence that potential confounders have been sufficiently adjusted for.

@@

A review of other available literature suggests that this observation (a harmful effect of lipid-regulating agents on vascular-related outcomes) is not unusual. Using a conventional epidemiological technique, a previous analysis also found an increased risk of coronary heart disease (analogous to the ischemic heart disease outcome used in this analysis) in those taking statins (`r estimate(1.31,1.04, 1.66, type = "HR")`).[@danaei2013b] Controlling for confounding by indication in that study through the use of a trial emulation analysis gave an estimate of 0.89 (95%CI: 0.73-1.09), a comparable though less precise estimate to that observed in RCTs of statin use (`r estimate(0.73,0.67,0.80,type="")`).[@taylor2013]

@@

However, the findings of this analysis are subject to several limitations in addition to the confounding by indication discussed above. There is a strong possibility of differential misclassification of dementia-related conditions based on the exposure.[@porta2014dictionary] As an illustrative example, those with memory complaints may be more likely to be classified as vascular dementia than Alzheimer's disease if their medical records contains prescriptions for lipid-regulating agents. Further, there is a potential for general non-differential misclassification of the outcome due to the varying positive predictive value of electronic health record code lists to identify dementia cases.[@wilkinson2018; @mcguinness2019validity]
<!-- though there is some encouraging evidence that a dedicated coding exercise increases the identification of dementia case in primary care database and does not add a substantial administrative burden.[@russell2013] -->

@@

The raw data supporting this analysis is not publicly available, as access to the CPRD data is controlled by a data monitoring committee. However, when data are not readily available, sharing the analysis code and summary statistics represents a way for readers to validate the findings.[@goldacre2019c]

@@

This chapter attempted to account for important sources of bias and provide a comparison with other available literature, as identified in the systematic review presented in Chapter \@ref(sys-rev-heading). However, there is a strong potential for uncontrolled confounding by indication and differential misclassification of the outcome on the basis of exposure, which raises questions about the findings, in particular the unexpected increase in risk of vascular dementia associated with statin use. The presence of bias is supported by our findings for the negative and positive control outcomes used, which provide some evidence of uncontrolled vascular confounders that may both prompt LRA prescription and increase risk of dementia outcomes with a vscular component. Future research using large scale electronic health records should aim to address these limitations, potentially by using a analytical design that more closely emulates a trial.[@danaei2013b]

>>>>

In this analysis, missing data were handled using a multiple imputation approach.[@sterne2009] Variables with missing observations were identified for inclusion in the imputation model. Nominal variables with missing values were modelled using multinomial logistic regression, while continuous variables were modelled using linear regression. As per best practice, all variables used in the analytic model, including the outcome, were included in the imputation model.[@moons2006] Using the MICE (Multiple Imputation by Chained Equations) command in STATA16, 20 imputed datasets were created.

@@

Using multiple imputation to handle missing data is an alternative to a "complete case" approach,[@pigott2001] where participants missing any covariate are dropped from the dataset. As a recommended sensitivity analysis,[@hughes2019] I preformed and compared the results of both methods, to investigate the impact of multiple imputation on the results.

@@

Despite observational analyses suggesting a link between statins and muscular pain (as opposed to more serious complications such as myopathy),[@selva-ocallaghan2018] systematic reviews of the adverse events of statin use[@collins2016] and N-of-1 trials explicitly exploring the association of statin use with muscle pain[@herrett2021] have found little evidence supporting an effect. This suggests that muscular backpain would be suitable for use as a negative control outcome in this analysis. Under this approach, if statin use is found to be associated with muscular backpain in this analysis, this suggests the presence of residual confounding and reduces my confidence in the results for the dementia outcomes.

Similarly, incident ischemic heart disease and Type 2 diabetes were included as a protective and harmful positive control outcome, respectively. The protective effect of lipid-lowering treatment, via statins, on the risk of ischemic heart disease is well-established, [@collins2016] while there is growing evidence for an increased risk of Type 2 diabetes with statin use.[@collins2016; @macedo2014; @smit2020] Similar to the negative outcome, if the analysis strategy can reproduce these known associations, this will provide evidence that potential confounders have been sufficiently adjusted for.

@@

A review of other available literature suggests that this observation (a harmful effect of lipid-regulating agents on vascular-related outcomes) is not unusual. Using a conventional epidemiological technique, a previous analysis also found an increased risk of coronary heart disease (analogous to the ischemic heart disease outcome used in this analysis) in those taking statins (`r estimate(1.31,1.04, 1.66, type = "HR")`).[@danaei2013] Controlling for confounding by indication in that study through the use of a trial emulation analysis gave an estimate of 0.89 (95%CI: 0.73-1.09), a comparable though less precise estimate to that observed in RCTs of statin use (`r estimate(0.73,0.67,0.80,type="")`).[@taylor2013]

@@

However, the findings of this analysis are subject to several limitations in addition to the confounding by indication discussed above. There is a strong possibility of differential misclassification of dementia-related conditions based on the exposure.[@porta2014] As an illustrative example, those with memory complaints may be more likely to be classified as vascular dementia than Alzheimer's disease if their medical records contains prescriptions for lipid-regulating agents. Further, there is a potential for general non-differential misclassification of the outcome due to the varying positive predictive value of electronic health record code lists to identify dementia cases.[@wilkinson2018; @mcguinness2019validity]
<!-- though there is some encouraging evidence that a dedicated coding exercise increases the identification of dementia case in primary care database and does not add a substantial administrative burden.[@russell2013] -->

@@

The raw data supporting this analysis is not publicly available, as access to the CPRD data is controlled by a data monitoring committee. However, when data are not readily available, sharing the analysis code and summary statistics represents a way for readers to validate the findings.[@goldacre2019]

@@

This chapter attempted to account for important sources of bias and provide a comparison with other available literature, as identified in the systematic review presented in Chapter \@ref(sys-rev-heading). However, there is a strong potential for uncontrolled confounding by indication and differential misclassification of the outcome on the basis of exposure, which raises questions about the findings, in particular the unexpected increase in risk of vascular dementia associated with statin use. The presence of bias is supported by our findings for the negative and positive control outcomes used, which provide some evidence of uncontrolled vascular confounders that may both prompt LRA prescription and increase risk of dementia outcomes with a vscular component. Future research using large scale electronic health records should aim to address these limitations, potentially by using a analytical design that more closely emulates a trial.[@danaei2013]

### 2021-11-23 06-Chp-IPD-Analysis.Rmd
Citations fixed!

Previous work has suggested a difference in the effect of lipids on dementia risk based on patient age and sex.[@ancelin2013; @mielke2010a] While the systematic review presented in Chapter \@ref(sys-rev-heading) did not identify strong evidence for an effect of sex or age at lipid measurement on the relationship between lipids and dementia. However, the number of included studies in the lipid fraction meta-analyses was small, due to both the relatively small number of studies reporting on the lipids exposure and the poor reporting of summary statistics of patient characteristics in those that did. Additionally, best practice guidance recommends against basing the decision to perform an IPD meta-analysis on the between-study heterogeneity in a summary data meta-analysis, as similar distributions of patient covariates across studies may mask a true effect.[@riley2020]

@@

Across all cohorts, data cleaning was performed in a similar manner, standardising to commonly named variables so that a single model could be applied using functional programming.[@wickham2016a] The advantage of this approach is that it reduces the likelihood of errors in model mis-specification if needing to change variables names from cohort to cohort. Following data cleaning, summary statistics for each data source were calculated and compared to publicly available statistics to ensure no errors were introduced in the data cleaning process, in line with best practice.[@levis2021]

@@

In order to investigate the interaction of patient-level characteristics (age and sex) with lipid levels, interaction terms for lipid-covariate terms were extracted and synthesised using a random effects meta-analysis.[@fisher2017a] To avoid ecological bias, where the between-study association does not reflect the within-study associations, cohorts where there was no within-study variation in the covariate of interest were excluded from the interaction analysis for that covariate.[@burke2017] As an example, it is impossible to estimate the impact of sex on the lipid~dementia relationship in a study which contains female participants. All interaction analysis were performed using the common-set-adjusted model (Model 1).

@@

A low response rate is not unexpected, given that a review of IPD studies published between 1987 and 2015 found that fewer than half managed to obtained data from greater than 80% of studies, and that in many cases, the exact percentage of studies for which data were obtained was not accurately reported.[@nevitt2017a] However, it is assumed that a ~10% response rate is at the lower end of the scale.

There are likely several likely reasons for this low response rate to requests for data access. In general terms, there are a several well-documented reasons why data are not made readily available, including concerns regarding participant privacy and "scooping" or "parasitic" behaviour, and a lack of trust between primary and secondary researchers.[@vanpanhuis2014] More specific to this analysis, individual participant data meta-analysis including studies other than randomised controlled trials have less success in obtaining individual participant data from studies.[@nevitt2017a] Additionally, while no evidence is available on whether the characteristics of the researcher requesting data access influences the response rate and eventual decision, there is the possibility that my position as a PhD student meant I was less likely to elicit the response rate than a well-known senior academic might. The timing of the requests for data access, coinciding with a global pandemic, may also have affected the response rate as researchers prioritised other COVID-related work.

>>>>

Previous work has suggested a difference in the effect of lipids on dementia risk based on patient age and sex.[@ancelin2013; @mielke2010] While the systematic review presented in Chapter \@ref(sys-rev-heading) did not identify strong evidence for an effect of sex or age at lipid measurement on the relationship between lipids and dementia. However, the number of included studies in the lipid fraction meta-analyses was small, due to both the relatively small number of studies reporting on the lipids exposure and the poor reporting of summary statistics of patient characteristics in those that did. Additionally, best practice guidance recommends against basing the decision to perform an IPD meta-analysis on the between-study heterogeneity in a summary data meta-analysis, as similar distributions of patient covariates across studies may mask a true effect.[@riley2020]

@@

Across all cohorts, data cleaning was performed in a similar manner, standardising to commonly named variables so that a single model could be applied using functional programming.[@wickham2016func] The advantage of this approach is that it reduces the likelihood of errors in model mis-specification if needing to change variables names from cohort to cohort. Following data cleaning, summary statistics for each data source were calculated and compared to publicly available statistics to ensure no errors were introduced in the data cleaning process, in line with best practice.[@levis2021]

@@

In order to investigate the interaction of patient-level characteristics (age and sex) with lipid levels, interaction terms for lipid-covariate terms were extracted and synthesised using a random effects meta-analysis.[@fisher2017] To avoid ecological bias, where the between-study association does not reflect the within-study associations, cohorts where there was no within-study variation in the covariate of interest were excluded from the interaction analysis for that covariate.[@burke2017] As an example, it is impossible to estimate the impact of sex on the lipid~dementia relationship in a study which contains female participants. All interaction analysis were performed using the common-set-adjusted model (Model 1).

@@

A low response rate is not unexpected, given that a review of IPD studies published between 1987 and 2015 found that fewer than half managed to obtained data from greater than 80% of studies, and that in many cases, the exact percentage of studies for which data were obtained was not accurately reported.[@nevitt2017] However, it is assumed that a ~10% response rate is at the lower end of the scale.

There are likely several likely reasons for this low response rate to requests for data access. In general terms, there are a several well-documented reasons why data are not made readily available, including concerns regarding participant privacy and "scooping" or "parasitic" behaviour, and a lack of trust between primary and secondary researchers.[@vanpanhuis2014] More specific to this analysis, individual participant data meta-analysis including studies other than randomised controlled trials have less success in obtaining individual participant data from studies.[@nevitt2017] Additionally, while no evidence is available on whether the characteristics of the researcher requesting data access influences the response rate and eventual decision, there is the possibility that my position as a PhD student meant I was less likely to elicit the response rate than a well-known senior academic might. The timing of the requests for data access, coinciding with a global pandemic, may also have affected the response rate as researchers prioritised other COVID-related work.

### 2021-11-23 07-Chp-Triangulation.Rmd
Citations fixed!

Once each individual result had been adjusted using this approach, I performed a random effects meta-analysis using the unadjusted and adjusted results. I also compared the adjusted results under the two scenarios of additive bias defined. results. I extracted the overall effect estimates, along with measures of heterogeneity in each case ($\tau^2$, $I^2$).[@higgins2003; @higgins2008b]

@@

Ideally, these prior distributions would be based on empirical data on the effect of different domains of bias/indirectness on a result. Most meta-epidemiological studies examining the effect of different methodological issues studied only randomised controlled trials,[@amer2021; @page2016] and the estimates of bias they produce may not hold for non-randomised study designs. However, the prior distributions defined here are broadly comparable to previously reported estimates of the impact of bias in non-randomised trials. For example, a previous simulation study estimated that, in the absence of true effect of LDL-c, ApoE4 would induce a RR of 1.09 with risk of dementia per 1-SD increase in LDL-c.[@iwagami2021a] This is comparable to the mean assigned to the prior distribution mapped to "Moderate" additive bias (N(0.08,0.05), Table \@ref(tab:priorsAdd-table)).

@@

A further issue in this regard is that rigourness of the tools may not be equivalent. This issue can be illustrated by the Ostergaard _et al._ study[@ostergaard2017] having low risk of bias across all domains and thus receiving a substantial proportion of the weight in the first case study (Figure \@ref(fig:fpLdlAd)). While the RoB suite of tools (RoB 2.0, ROBINS-I, ROBINS-E) were developed by an expert working group and have been widely used, the Mamluk _et al._ tool[@mamluk2020] used to assess risk of bias in Mendelian randomisation studies was designed by the authors for use in their own review. If the Mamluk tool is failing to adequately assess bias in Mendelian randomisation studies, then any analysis based on its assessments is likely to be subject to bias. This observation is particularly problematic given the default position of the ROBINS-I and ROBINS-E tool to require a "Moderate" judgement in the confounding domain. Under these conditions, it is possible that a well performed cohort study (low risk of bias across everything other that bias due to confounding) will be down-weighted versus a potentially poor MR study.

>>>>

Once each individual result had been adjusted using this approach, I performed a random effects meta-analysis using the unadjusted and adjusted results. I also compared the adjusted results under the two scenarios of additive bias defined. results. I extracted the overall effect estimates, along with measures of heterogeneity in each case ($\tau^2$, $I^2$).[@higgins2003; @higgins2008]

@@

Ideally, these prior distributions would be based on empirical data on the effect of different domains of bias/indirectness on a result. Most meta-epidemiological studies examining the effect of different methodological issues studied only randomised controlled trials,[@amer2021; @page2016] and the estimates of bias they produce may not hold for non-randomised study designs. However, the prior distributions defined here are broadly comparable to previously reported estimates of the impact of bias in non-randomised trials. For example, a previous simulation study estimated that, in the absence of true effect of LDL-c, ApoE4 would induce a RR of 1.09 with risk of dementia per 1-SD increase in LDL-c.[@iwagami2021] This is comparable to the mean assigned to the prior distribution mapped to "Moderate" additive bias (N(0.08,0.05), Table \@ref(tab:priorsAdd-table)).

@@

A further issue in this regard is that rigourness of the tools may not be equivalent. This issue can be illustrated by the Ostergaard _et al._ study[@ostergaard2015] having low risk of bias across all domains and thus receiving a substantial proportion of the weight in the first case study (Figure \@ref(fig:fpLdlAd)). While the RoB suite of tools (RoB 2.0, ROBINS-I, ROBINS-E) were developed by an expert working group and have been widely used, the Mamluk _et al._ tool[@mamluk2020] used to assess risk of bias in Mendelian randomisation studies was designed by the authors for use in their own review. If the Mamluk tool is failing to adequately assess bias in Mendelian randomisation studies, then any analysis based on its assessments is likely to be subject to bias. This observation is particularly problematic given the default position of the ROBINS-I and ROBINS-E tool to require a "Moderate" judgement in the confounding domain. Under these conditions, it is possible that a well performed cohort study (low risk of bias across everything other that bias due to confounding) will be down-weighted versus a potentially poor MR study.
